Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-10-15 manuscript/coverletter/collabra_response_2.tex
futzing for resubmit

	We have added bolded text to our README linking to these different ways of viewing the links in a hyperlinked format.

>>>>

	We have changed the link to the paper for the pre-registration to point to the clickable links on OSF. We have added text to our README linking to these different ways of viewing the links in a hyperlinked format.

### 2024-10-15 manuscript/manuscript.Rmd
futzing for resubmit

---
title: "Estimating the replicability of psychology experiments after an initial failure to replicate"
author:
  - Anonymous
# author:
#   - name: Veronica Boyce
#     affiliation: Psych
#     footnote:
#       - corresp
#   - name: Ben Prystawski
#     affiliation: Psych
#   - name: Adani B. Abutto
#     affiliation: Psych
#   - name: Emily M. Chen
#     affiliation: Psych
#   - name: Ziwen Chen
#     affiliation: GSB
#   - name: Howard Chiu
#     affiliation: Edu
#   - name: Irmak Ergin
#     affiliation: Psych
#   - name: Anmol Gupta
#     affiliation: Psych
#   - name: Chuqi Hu
#     affiliation: Symsys
#   - name: Bendix Kemmann
#     affiliation: Philosophy
#   - name: Nastasia Klevak
#     affiliation: Psych
#   - name: Verity Y. Q. Lua
#     affiliation: Psych
#   - name: Mateus M. Mazzaferro
#     affiliation: Edu
#   - name: Khaing Mon
#     affiliation: Symsys
#   - name: Dan Ogunbamowo
#     affiliation: Psych
#   - name: Alexander Pereira
#     affiliation: [Psych, Philosophy]
#   - name: Jordan Troutman
#     affiliation: CS
#   - name: Sarah Tung
#     affiliation: Psych
#   - name: Raphael Uricher
#     affiliation: Psych
#   - name: Michael C. Frank
#     affiliation: Psych
# address:
#   - code: Psych
#     address: Department of Psychology, Stanford University
#   - code: GSB
#     address: Graduate School of Business, Stanford University
#   - code: Edu
#     address: Graduate School of Education, Stanford University
#   - code: Symsys
#     address: Symbolic Systems Program, Stanford University
#   - code: Philosophy
#     address: Department of Philosophy, Stanford University
#   - code: CS
#     address: Department of Computer Science, Stanford University
#footnote:
#  - code: corresp
#    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251rescue.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: apa7.csl # Use any CSL
nocite: |
  @krauss2003, @ngo2019, @todd2016, @jara-ettinger2022, @porter2016, @birch2007, @child2018, @chou2016, @craig2014, @gong2019, @haimovitz2016, @hopkins2016, @paxton2012, @payne2008, @schechtman2010, @tarampi2016, @yeshurun2003
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt
header-includes:
 - \usepackage{tikz}
 - \usetikzlibrary{positioning,chains}
 - \usepackage{setspace}\singlespacing
 - \renewcommand{\textfraction}{0.00}
 - \renewcommand{\topfraction}{1}
 - \renewcommand{\bottomfraction}{1}
 - \renewcommand{\floatpagefraction}{1}
 - \setcounter{topnumber}{3}
 - \setcounter{bottomnumber}{3}
 - \setcounter{totalnumber}{4}
editor_options: 
  markdown: 
    wrap: sentence
---

@@

Our analysis plan was pre-registered after students had selected projects, but before final data collection on the projects.
Each project was also individually pre-registered by the student conducting it.
The overall analysis is pre-registered at <https://osf.io/5qz7v>, individual pre-registrations are linked from <https://osf.io/vkbfw>.
We note one deviation from our pre-registration here: we pre-registered visual comparisons between original, first replication, and rescue projects using prediction intervals.
Prediction intervals depend on both the original effect size and variance and the variance of the comparison (replication or rescue) study [@patil2016].
Thus we cannot show a single prediction interval for the original study, but would have to show a prediction interval between each pair of studies.
We thought this analysis would not offer additional clarity and hence we did not perform it.

@@

Students were required to think of reasons why the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (<https://osf.io/dfwk2>).
In general, we encouraged students to add manipulation checks as appropriate and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of @paxton2012 switched from using the Cognitive Reflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2 which is used less but attempts to measure the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants use a dropdown menu to select left or right for each turn.
The rescue removed the unwieldy dropdown menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

>>>>

---
title: "Estimating the replicability of psychology experiments after an initial failure to replicate"
author:
  - name: Veronica Boyce
    affiliation: Psych
    footnote:
     - corresp
  - name: Ben Prystawski
    affiliation: Psych
  - name: Adani B. Abutto
    affiliation: Psych
  - name: Emily M. Chen
    affiliation: Psych
  - name: Ziwen Chen
    affiliation: GSB
  - name: Howard Chiu
    affiliation: Edu
  - name: Irmak Ergin
    affiliation: Psych
  - name: Anmol Gupta
    affiliation: Psych
  - name: Chuqi Hu
    affiliation: Symsys
  - name: Bendix Kemmann
    affiliation: Philosophy
  - name: Nastasia Klevak
    affiliation: Psych
  - name: Verity Y. Q. Lua
    affiliation: Psych
  - name: Mateus M. Mazzaferro
    affiliation: Edu
  - name: Khaing Mon
    affiliation: Symsys
  - name: Dan Ogunbamowo
    affiliation: Psych
  - name: Alexander Pereira
    affiliation: [Psych, Philosophy]
  - name: Jordan Troutman
    affiliation: CS
  - name: Sarah Tung
    affiliation: Psych
  - name: Raphael Uricher
    affiliation: Psych
  - name: Michael C. Frank
    affiliation: Psych
address:
  - code: Psych
    address: Department of Psychology, Stanford University
  - code: GSB
    address: Graduate School of Business, Stanford University
  - code: Edu
    address: Graduate School of Education, Stanford University
  - code: Symsys
    address: Symbolic Systems Program, Stanford University
  - code: Philosophy
    address: Department of Philosophy, Stanford University
  - code: CS
    address: Department of Computer Science, Stanford University
footnote:
 - code: corresp
   text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251rescue.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: apa7.csl # Use any CSL
nocite: |
  @krauss2003, @ngo2019, @todd2016, @jara-ettinger2022, @porter2016, @birch2007, @child2018, @chou2016, @craig2014, @gong2019, @haimovitz2016, @hopkins2016, @paxton2012, @payne2008, @schechtman2010, @tarampi2016, @yeshurun2003
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt
header-includes:
 - \usepackage{tikz}
 - \usetikzlibrary{positioning,chains}
 - \usepackage{setspace}\singlespacing
 - \renewcommand{\textfraction}{0.00}
 - \renewcommand{\topfraction}{1}
 - \renewcommand{\bottomfraction}{1}
 - \renewcommand{\floatpagefraction}{1}
 - \setcounter{topnumber}{3}
 - \setcounter{bottomnumber}{3}
 - \setcounter{totalnumber}{4}
editor_options: 
  markdown: 
    wrap: sentence
---

@@

Our analysis plan was pre-registered after students had selected projects, but before final data collection on the projects.
Each project was also individually pre-registered by the student conducting it.
The overall analysis is pre-registered at <https://osf.io/5qz7v>, individual pre-registrations are linked from <https://osf.io/r2d6q>.
We note one deviation from our pre-registration here: we pre-registered visual comparisons between original, first replication, and rescue projects using prediction intervals.
Prediction intervals depend on both the original effect size and variance and the variance of the comparison (replication or rescue) study [@patil2016].
Thus we cannot show a single prediction interval for the original study, but would have to show a prediction interval between each pair of studies.
We thought this analysis would not offer additional clarity and hence we did not perform it.

@@

Students were required to think of reasons why the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (<https://osf.io/dfwk2>).
In general, we encouraged students to add manipulation checks as appropriate and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of @paxton2012 switched from using the Cognitive Reflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2, which is used less but attempts to measure the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants use a dropdown menu to select left or right for each turn.
The rescue removed the unwieldy dropdown menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.
